Giant Low Surface Brightness Galaxies As Merger Remnants

### 2016-07-11 glsb.tex
add califa paper
add more discussion

For each galaxy, observations were taken at three unique dither positions to provide nearly complete spatial coverage.  Individual science exposures were 30 minutes.  Malin 2 had 10 exposures per dither position (11 exposures for dither position 1), for a total of 15 hours of exposure time.  Observations of UGC 06614 had alignment issues between observing runs.  This resulted in a deep stack of 6 exposures, and a shallow stack with only 2 exposures.  With only 2 exposures, cosmic ray rejection becomes difficult, therefore, we use these shallow frames when looking at the surface brightness profiles, but use only the aligned deep co-add when analyzing the stellar kinematics and stellar populations. Twilight flat field frames, bias frames, and wavelength calibration Hg and Cd lamps were taken during the day.  Flux standard stars were observed in twilight.

@@

@@

An accretion formation scenario does the best job explaining all the observations.  While a collisional ring could fade to a GLSB, the stellar population should be relatively young, and one might not expect a steep metallicity gradient.  Also, there are no obvious galaxies near GLSBs to act as progenitor colliders.

It also seems unlikely that GLSBs are just extreme versions of normal disk galaxies \citep{Das13}.  It's hard to imagine how such a large massive central spheroid could be formed without also disrupting the surrounding disk.  While the old stellar populations in the disks of Malin~2 and UGC~6614 are evidence against the \citet{Mapelli08} formation scenario, we have not addressed what collisional ring galaxies do evolve into.  Since photometric surveys are often surface brightness limited, it is possible a GLSB system will be found in the future that is the decendent of a collisional ring system.

@@

\citet{Kasparova14} present one of the most in depth observational studies of Malin~2.  Mass-modeling the system, they find that a huge isothermal core with $r_c=27.3$ kpc.  \citet{Kasparova14} also measure gas metallicities and find the metallicity drops from solar in the center of the system to half-solar at $r=30$\ kpc.  It is probably foolish to directly compare gas and stellar metallicities, but it is interesting that the gas metallicity gradient appears to be much lower than the stellar gradient.  This is probably consistent with a low metallicity gas-rich galaxy merging with an elliptical system that has much higher stellar metallicity, although a detailed simulation would be heklpful in confirming what the final expected gas metallicity gradient should be.

>>>>

For each galaxy, observations were taken at three unique dither positions to provide nearly complete spatial coverage.  Individual science exposures were 30 minutes.  Malin~2 had 10 exposures per dither position (11 exposures for dither position 1), for a total of 15 hours of exposure time.  Observations of UGC 06614 had alignment issues between observing runs.  This resulted in a deep stack of 6 exposures, and a shallow stack with only 2 exposures.  With only 2 exposures, cosmic ray rejection becomes difficult, therefore, we use these shallow frames when looking at the surface brightness profiles, but use only the aligned deep co-add when analyzing the stellar kinematics and stellar populations. Twilight flat field frames, bias frames, and wavelength calibration Hg and Cd lamps were taken during the day.  Flux standard stars were observed in twilight.

@@

XXX-organize this up a little more.

@@

An accretion formation scenario does the best job explaining all the observations.  While a collisional ring could fade to a GLSB, the stellar population should be relatively young, and one would not necissarily expect a steep metallicity gradient.  Also, there are no obvious galaxies near GLSBs to act as progenitor colliders.

It also seems unlikely that GLSBs are just extreme versions of normal disk galaxies \citep{Das13}.  It's hard to imagine how such a large massive central spheroid could be formed without also disrupting the surrounding disk.  While the old stellar populations in the disks of Malin~2 and UGC~6614 are evidence against the \citet{Mapelli08} formation scenario, we have not addressed what collisional ring galaxies do evolve into.  Since photometric surveys are often surface brightness limited, it is possible a GLSB system will be found in the future that is the descendant of a collisional ring system.

@@

\citet{Sanchez14} measure the Oxygen abundance from gas emission lines in a large sample of spiral galaxies and find a characteristic slope of $\alpha_{\rm{O/H}} = -0.1$\ dex $r_e^{-1}$ between 0.3 and 2 effective radii.  The metallicity gradients we observe in Malin~2 and UGC~6614 are factors of 2-4 steeper than the typical spiral. Also, of the 193 galaxies in \citet{Sanchez14} only four have a slope steeper than -0.3 \ dex $r_e^{-1}$ \citep[Figure~6]{Sanchez14}.  In their Figure~9, they combine the data to show a typical gas metallicity curve that drops linearly until $r_e\sim2$ and then flattens to a constant value.  This shape is very different from the step function change we observe in the GLSBs.

% malin 2: R_e = 7.17"
% UGC : 7.3"

\citet{Kasparova14} present one of the most in depth observational studies of Malin~2.  Mass-modeling the system, they find that a huge isothermal core with $r_c=27.3$ kpc.  \citet{Kasparova14} also measure gas metallicities and find the metallicity drops from solar in the center of the system to half-solar at $r=30$\ kpc.  It is probably foolish to directly compare gas and stellar metallicities, but it is interesting that the gas metallicity gradient appears to be much lower than the stellar gradient.  This is probably consistent with a low metallicity gas-rich galaxy merging with an elliptical system that has much higher stellar metallicity, although a detailed simulation would be helpful in confirming what the final expected gas metallicity gradient should be.
